---
title: "Other Examples"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Other Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Example 1 

Time-independent dataset, when the dataset can be saved and loaded as a whole: Fitting a Cox model for a time-independent dataset with 50 variables and 100,000 samples (detailed version of Key Example 1).

```{r, eval=FALSE}
##### Generating a time-independent dataset #####
set.seed(1)
N <- 1e5
p.x <- 50
K <- 100
n <- N / K
cor <- 0.2
bb <- c(rep(0.4, 4), rep(0.2, 4), rep(0.1, 4), rep(0.05, 4))
beta0 <- c(1, bb, rep(0, p.x - length(bb)))
dat.mat0 <- as.data.frame(SIM.FUN(N, p.x = p.x, cor = cor, family = "Cox", beta0 = beta0))
dat.mat0[, "strat"] <- rep(1:20, each = N / 20)

## A Cox model, without stratification of baseline hazard: Surv(u,delta)~V3+V4+.....+V52
# Example option 1: Using 1 core for computation, dividing the dataset to 10 chunks, using 2 iterations of one-step estimator for update
mod <- dcalasso(as.formula(paste0("Surv(u,delta)~", paste(paste0("V", 3:52), collapse = "+"))),
  family = "cox.ph", data = dat.mat0,
  K = 10, iter.os = 2
)
sum.mod <- summary(mod)
print(sum.mod, unpen = T)
plot(mod)

# Performing model prediction
pred.link <- predict(mod, newdata = dat.mat0)
pred.term <- predict(mod, newdata = dat.mat0, type = "terms")
pred.response <- predict(mod, newdata = dat.mat0, type = "response")


# Example option 2: Same model as above: Using 2 cores for parallel computation, dividing the dataset to 10 chunks, using 4 iterations of one-step estimator for update
modp <- dcalasso(as.formula(paste0("Surv(u,delta)~", paste(paste0("V", 3:52), collapse = "+"))),
  family = "cox.ph", data = dat.mat0,
  K = 10, iter.os = 4, ncores = 2
)
sum.modp <- summary(modp)
print(sum.modp, unpen = T)
plot(modp)

# Compare the divide and conquer method unpenalized model fit against a standard Cox fit using the whole dataset
std <- coxph(as.formula(paste0("Surv(u,delta)~", paste(paste0("V", 3:52), collapse = "+"))),
  data = dat.mat0
)

plot(mod$coefficients.unpen, std$coefficients)
plot(modp$coefficients.unpen, std$coefficients)


## A Cox model, with stratification of baseline hazard by "strat": Surv(u,delta)~strata(strat)+V3+V4+.....+V52
# Example option 1: Using 1 core for computation, dividing the dataset to 10 chunks, using 2 iterations of one-step estimator for update
mod <- dcalasso(as.formula(paste0("Surv(u,delta)~strata(strat)+", paste(paste0("V", 3:52), collapse = "+"))),
  family = "cox.ph", data = dat.mat0,
  K = 10, iter.os = 2
)
sum.mod <- summary(mod)
print(sum.mod, unpen = T)
plot(mod)

# Example option 2: Same model as above: Using 2 cores for parallel computation, dividing the dataset to 10 chunks, using 4 iterations of one-step estimator for update
modp <- dcalasso(as.formula(paste0("Surv(u,delta)~strata(strat)+", paste(paste0("V", 3:52), collapse = "+"))),
  family = "cox.ph", data = dat.mat0,
  K = 10, iter.os = 2, ncores = 2
)
sum.modp <- summary(modp)
print(sum.modp, unpen = T)
plot(modp)

# Compare the divide and conquer method unpenalized model fit against a standard Cox fit using the whole dataset
std <- coxph(as.formula(paste0("Surv(u,delta)~strata(strat)+", paste(paste0("V", 3:52), collapse = "+"))),
  data = dat.mat0
)


plot(mod$coefficients.unpen, std$coefficients)
plot(modp$coefficients.unpen, std$coefficients)
```


## Example 2 

Time-independent dataset, when the dataset are saved in multiple files: Fitting a Cox model for a time-independent dataset with 50 variables and 100,000 samples.

```{r, eval=FALSE}
##### Generating a time-independent dataset, saving the data into 10 separate files #####
set.seed(1)
N <- 1e5
p.x <- 50
K <- 100
n <- N / K
cor <- 0.2
bb <- c(rep(0.4, 4), rep(0.2, 4), rep(0.1, 4), rep(0.05, 4))
beta0 <- c(1, bb, rep(0, p.x - length(bb)))
dat.mat0 <- as.data.frame(SIM.FUN(N, p.x = p.x, cor = cor, family = "Cox", beta0 = beta0))
dat.mat0[, "strat"] <- rep(1:20, each = N / 20)
dir <- "C:/"
ll <- split(1:N, factor(1:10))
for (kk in 1:10) {
  df <- dat.mat0[ll[[kk]], ]
  saveRDS(df, file = paste0(dir, "dataTI", kk, ".rds"))
}

## A Cox model, without stratification of baseline hazard: Surv(u,delta)~V3+V4+.....+V52
# Example option 1: Using 1 core for computation, data.rds specifies that the data are saved into 10 files [each time the RAM will only contain data from 1 file], using 2 iterations of one-step estimator for update
mod <- dcalasso(as.formula(paste0("Surv(u,delta)~", paste(paste0("V", 3:52), collapse = "+"))),
  family = "cox.ph",
  data.rds = paste0(dir, "dataTI", 1:10, ".rds"), iter.os = 2
)
sum.mod <- summary(mod)
print(sum.mod, unpen = T)
plot(mod)

# Example option 2: Using 2 core for parallel computation, data.rds specifies that the data are saved into 10 files [each time the RAM will only contain data from 1 file], using 2 iterations of one-step estimator for update
modp <- dcalasso(as.formula(paste0("Surv(u,delta)~", paste(paste0("V", 3:52), collapse = "+"))),
  family = "cox.ph",
  data.rds = paste0(dir, "dataTI", 1:10, ".rds"), iter.os = 2, ncores = 2
)
sum.modp <- summary(modp)
print(sum.modp, unpen = T)
plot(modp)

# Compare the divide and conquer method unpenalized model fit against a standard Cox fit using the whole dataset
std <- coxph(as.formula(paste0("Surv(u,delta)~", paste(paste0("V", 3:52), collapse = "+"))),
  data = dat.mat0
)

plot(mod$coefficients.unpen, std$coefficients)
plot(modp$coefficients.unpen, std$coefficients)


## A Cox model, with stratification of baseline hazard by "strat": Surv(u,delta)~strata(strat)+V3+V4+.....+V52
# Example option 1: Using 1 core for computation, loading data from 10 separate files (imply K=10) [each time the RAM will only contain data from 1 file], using 2 iterations of one-step estimator for update
mod <- dcalasso(as.formula(paste0("Surv(u,delta)~strata(strat)+", paste(paste0("V", 3:52), collapse = "+"))),
  family = "cox.ph",
  data.rds = paste0(dir, "dataTI", 1:10, ".rds"), iter.os = 2
)
sum.mod <- summary(mod)
print(sum.mod, unpen = T)
plot(mod)

# Example option 2: Using 2 core for parallel computation, data.rds specifies that the data are saved into 10 files [each time the RAM will only contain data from 1 file], using 2 iterations of one-step estimator for update
modp <- dcalasso(as.formula(paste0("Surv(u,delta)~strata(strat)+", paste(paste0("V", 3:52), collapse = "+"))),
  family = "cox.ph",
  data.rds = paste0(dir, "dataTI", 1:10, ".rds"), iter.os = 2, ncores = 2
)
sum.modp <- summary(modp)
print(sum.modp, unpen = T)
plot(modp)

# Compare the divide and conquer method unpenalized model fit against a standard Cox fit using the whole dataset
std <- coxph(as.formula(paste0("Surv(u,delta)~strata(strat)+", paste(paste0("V", 3:52), collapse = "+"))),
  data = dat.mat0
)

plot(mod$coefficients.unpen, std$coefficients)
plot(modp$coefficients.unpen, std$coefficients)
```


## Example 3 

Time-dependent dataset, when the dataset can be saved and loaded as a whole: Fitting a Cox model for a dataset with 50 time-dependent variables, 50 additional time-independent variables, and 100,000 samples (detailed version of Key Example 2).

```{r, eval=FALSE}
########### Generating a time-dependent dataset ####################
set.seed(1)
n.subject <- 1e5
p.ti <- 50
p.tv <- 50
K <- 20
n <- n.subject / K
cor <- 0.2
lambda.grid <- 10^seq(-10, 3, 0.01)
beta0.ti <- NULL
beta0.tv <- NULL
dat.mat0 <- as.data.frame(SIM.FUN.TVC(p.ti, p.tv, n.subject, cor, beta0.ti, beta0.tv))
dat.mat0[, "strat"] <- dat.mat0[, dim(dat.mat0)[2]] %% (n.subject / 20)
dat.mat0 <- dat.mat0[, -(dim(dat.mat0)[2] - 1)]

## A time-dependent Cox model, without stratification of baseline hazard: Surv(t0,t1,status)~V4+V5+...+V103
# Example option 1: Using 1 core for computation, dividing the data into 10 chunks, using 2 iterations of one-step estimator for update
mod <- dcalasso(as.formula(paste0("Surv(t0,t1,status)~", paste(paste0("V", 4:103), collapse = "+"))),
  family = "cox.ph", data = dat.mat0,
  K = 10, iter.os = 2
)
sum.mod <- summary(mod)
print(sum.mod, unpen = T)
plot(mod)

# Example option 2: Using 2 cores for parallel computation, dividing the data into 10 chunks, using 2 iterations of one-step estimator for update
modp <- dcalasso(as.formula(paste0("Surv(t0,t1,status)~", paste(paste0("V", 4:103), collapse = "+"))),
  family = "cox.ph", data = dat.mat0,
  K = 10, iter.os = 2, ncores = 2
)
sum.modp <- summary(modp)
print(sum.modp, unpen = T)
plot(modp)

# Compare the divide and conquer method unpenalized model fit against a standard Cox fit using the whole dataset
std <- coxph(as.formula(paste0("Surv(t0,t1,status)~", paste(paste0("V", 4:103),
  collapse = "+"
))),
data = dat.mat0
)
plot(mod$coefficients.unpen, std$coefficients)
plot(modp$coefficients.unpen, mod$coefficients.unpen)

## A time-dependent Cox model, with stratification of baseline hazard: Surv(t0,t1,status)~strata(strat)+V4+V5+...+V103
# Example option 1: Using 1 core for computation, dividing the data into 10 chunks, using 4 iterations of one-step estimator for update
mod <- dcalasso(as.formula(paste0("Surv(t0,t1,status)~strata(strat)+", paste(paste0("V", 4:103), collapse = "+"))),
  family = "cox.ph", data = dat.mat0,
  K = 10, iter.os = 4
)
sum.mod <- summary(mod)
print(sum.mod, unpen = T)
plot(mod)

# Example option 2: Using 2 cores for parallel computation, dividing the data into 10 chunks, using 2 iterations of one-step estimator for update
modp <- dcalasso(as.formula(paste0("Surv(t0,t1,status)~strata(strat)+", paste(paste0("V", 4:103), collapse = "+"))),
  family = "cox.ph", data = dat.mat0,
  K = 10, iter.os = 4, ncores = 2
)
sum.modp <- summary(modp)
print(sum.modp, unpen = T)
plot(modp)

# Compare the divide and conquer method unpenalized model fit against a standard Cox fit using the whole dataset
std <- coxph(as.formula(paste0("Surv(t0,t1,status)~strata(strat)+", paste(paste0("V", 4:103),
  collapse = "+"
))),
data = dat.mat0
)
plot(mod$coefficients.unpen, std$coefficients)
plot(modp$coefficients.unpen, mod$coefficients.unpen)
```

## Example 4 

Time-dependent dataset, when the dataset are saved in multiple files: Fitting a Cox model for a dataset with 50 time-dependent variables, 50 additional time-independent variables, and 100,000 samples.

```{r, eval=FALSE}
########### Generating a time-dependent dataset, saving the dataset to 10 separate files ####################
set.seed(1)
n.subject <- 1e5
p.ti <- 50
p.tv <- 50
K <- 20
n <- n.subject / K
cor <- 0.2
lambda.grid <- 10^seq(-10, 3, 0.01)
beta0.ti <- NULL
beta0.tv <- NULL
dat.mat0 <- as.data.frame(SIM.FUN.TVC(p.ti, p.tv, n.subject, cor, beta0.ti, beta0.tv))
dat.mat0[, "strat"] <- dat.mat0[, dim(dat.mat0)[2]] %% (n.subject / 20)
dat.mat0 <- dat.mat0[, -(dim(dat.mat0)[2] - 1)]
ll <- split(1:dim(dat.mat0)[1], factor(1:10))
for (kk in 1:10) {
  df <- dat.mat0[ll[[kk]], ]
  saveRDS(df, file = paste0(dir, "dataTV", kk, ".rds"))
}


## A time-dependent Cox model, without stratification of baseline hazard: Surv(t0,t1,status)~V4+V5+...+V103
# Example option 1: Using 1 core for computation, loading dataset from 10 chunks [each time the RAM will only contain data from 1 file], using 2 iterations of one-step estimator for update
mod <- dcalasso(as.formula(paste0("Surv(t0,t1,status)~", paste(paste0("V", 4:103), collapse = "+"))),
  family = "cox.ph",
  data.rds = paste0(dir, "dataTV", 1:10, ".rds"), iter.os = 2
)
sum.mod <- summary(mod)
print(sum.mod, unpen = T)
plot(mod)

# Example option 2: Using 2 cores for parallel computation, loading dataset from 10 chunks [each time the RAM will only contain data from 1 file], using 2 iterations of one-step estimator for update
modp <- dcalasso(as.formula(paste0("Surv(t0,t1,status)~", paste(paste0("V", 4:103), collapse = "+"))),
  family = "cox.ph",
  data.rds = paste0(dir, "dataTV", 1:10, ".rds"), iter.os = 2, ncores = 2
)
sum.modp <- summary(modp)
print(sum.modp, unpen = T)
plot(modp)

# Compare the divide and conquer method unpenalized model fit against a standard Cox fit using the whole dataset
std <- coxph(as.formula(paste0("Surv(t0,t1,status)~", paste(paste0("V", 4:103),
  collapse = "+"
))),
data = dat.mat0
)
plot(mod$coefficients.unpen, std$coefficients)
plot(modp$coefficients.unpen, mod$coefficients.unpen)

## A time-dependent Cox model, with stratification of baseline hazard: Surv(t0,t1,status)~strata(strat)+V4+V5+...+V103
# Example option 1: Using 1 core for computation, loading dataset from 10 chunks [each time the RAM will only contain data from 1 file], using 4 iterations of one-step estimator for update
mod <- dcalasso(as.formula(paste0("Surv(t0,t1,status)~strata(strat)+", paste(paste0("V", 4:103), collapse = "+"))),
  family = "cox.ph",
  data.rds = paste0(dir, "dataTV", 1:10, ".rds"), iter.os = 4
)
sum.mod <- summary(mod)
print(sum.mod, unpen = T)
plot(mod)

# Example option 2: Using 2 cores for parallel computation, loading dataset from 10 chunks [each time the RAM will only contain data from 1 file], using 4 iterations of one-step estimator for update
modp <- dcalasso(as.formula(paste0("Surv(t0,t1,status)~strata(strat)+", paste(paste0("V", 4:103), collapse = "+"))),
  family = "cox.ph",
  data.rds = paste0(dir, "dataTV", 1:10, ".rds"), K = 10, iter.os = 4, ncores = 2
)
sum.modp <- summary(modp)
print(sum.modp, unpen = T)
plot(modp)

# Compare the divide and conquer method unpenalized model fit against a standard Cox fit using the whole dataset
std <- coxph(as.formula(paste0("Surv(t0,t1,status)~strata(strat)+", paste(paste0("V", 4:103),
  collapse = "+"
))),
data = dat.mat0
)
plot(mod$coefficients.unpen, std$coefficients)
plot(modp$coefficients.unpen, mod$coefficients.unpen)
```
